Incremental tactics

### 2012-06-08 doc/papers/LFMTP2012/paper.tex
paper: fix & improve system rules
* only thread Delta and not Repo
* add syntactic cat. of atomic types, strengthen rules
* repo contains atomic terms

\begin{align*}
  K &\gequal
  \prd x A K \gor
  \type &
  \text{Kind}\\
  A &\gequal
  \prd x A A \gor
  \app {\cst a} S &
  \text{Type family} \\
  M &\gequal
  \lam x M \gor
  F &
  \text{Canonical object} \\
  F &\gequal \app H S
  \gor
  \smeta X \sigma &
  \text{Atomic object} \\
  H &\gequal
  \var x \gor
  \cst c &
  \text{Head}\\
  S &\gequal
  \spinenil \gor
  \spinecons M S &
  \text{Spine}\\
  \sigma &\gequal
  \msubstnil \gor
  \msubstcons \sigma x M &
  \text{Parallel substitution} \\
  \Gamma &\gequal
  \envnil \gor
  \envcons \Gamma x A &
  \text{Environment}
\end{align*}

@@

A \emph{repository} $\mr$ is a finite map from metavariable to
judgements of object well-typing, along with a distinguished
metavariable, corresponding to the \emph{tip} of the term contained in
it:
$$ \mr\;:\;(\meta X \mapsto (\Gamma\vdash\tp M A)), \smeta X
\sigma $$
The algorithm will maintain the invariant that a repository
$\mr$ is \emph{acyclic}, that is for all $\meta X$ in $\mr$, the
fixpoint of $\FMV{\meta X}$ does not contain $\meta X$. If $\mr =
(\Delta, X[\sigma])$, we write $\hat {\mr}$ for $X[\sigma]$, abusively
$\mr$ for $\Delta$ and $\mr (\smeta X \sigma)$ for $\gsubst M\sigma$
if $\Delta(\meta X) = (\Gamma\vdash M : A)$.

@@

  \fbox{$\jm\mr\Gamma M A \mmr {M'}$}
  \qquad
  Canonical object

  \begin{mathpar}
    % Lambda
    \infer{
      \jm\mr{\envcons\Gamma x A} M {A'} \mmr {M'}
    }{
      \jm\mr\Gamma {\lam x M} {\prd x A {A'}} \mmr {\lam x M'}
    }

    % App
    \infer{
      \jf\mr\Gamma F \mmr {F'} {A'}
       \and
      \jea\mmr\Gamma {\app{\cst a} S} {A'}
    }{
      \jm\mr\Gamma F {\app{\cst a} S} \mmr {F'}
    }
    % TODO commenter type ground
  \end{mathpar}

  \fbox{$\jf\mr\Gamma F \mmr F A$}
  \qquad
  {Atomic object}

@@

    % Non-sliceable
    \infer{
      \jh\mr\Gamma H A \annnonsliceable
      \and
      \jl\mr\Gamma A S \mmr {S'} {A'}
    }{
      \jf\mr\Gamma {\app H S} \mmr {\app H {S'}} {A'}
    }

    % Meta
    \infer{
      \mr(\meta X) = (\Gamma'\vdash M:A)
      \and
      \js\mr\Gamma \sigma {\Gamma'} \mmr {\sigma'}
    }{
      \jf\mr\Gamma {\smeta X \sigma} \mmr {\smeta X {\sigma'}} {\gsubst{A}\sigma'}
    }

    % Sliceable
    \infer{
      \jh\mr\Gamma H A \annsliceable
      \and
      \jl\mr\Gamma A S \mmr {S'} {A'}
      \and
      \stren{\Gamma, \app H {S'}}=\Gamma'
    }{
      \jf\mr\Gamma {\app H S} {\mapadd\mmr{\meta X}{(\Gamma'\vdash
          {\app H {S'}}:A)}} {\smeta X{\msubstid{\Gamma'}}} {A'}
    }
    \quad\text{($\meta X$ fresh in $\mr$)}

    % Defined
    \infer{
      \jh\mr\Gamma H A {\anndefined T}
      \and
      \jl\mr\Gamma A S {\mmr} {S'} {A'}
      \and
      \jcc\mr\Gamma A T S \mmmr F
      \and
      \jm\mmmr\Gamma F {A'} {\repo {R'''} } {F'}
    }{
      \jf\mr\Gamma {\app H S} {\repo{R'''}} {F'} {A'}
    }

@@

  \fbox{$\jh\mr\Gamma H A L$}
  \qquad
  {Head}

  \begin{mathpar}
    % constant
    \infer{
      \Sigma(\cst c) = A; L
    }{
      \jh\mr\Gamma {\cst c} A L
    }

    % variable
    \infer{
      \Gamma(\var x) = A
    }{
      \jh\mr\Gamma {\var x} A \annnonsliceable
    }
  \end{mathpar}

  \fbox{$\jl\mr\Gamma A S \mmr{S'}{A'}$}
  \qquad
  {Spine}

  \begin{mathpar}
    \infer{
      \jm\mr\Gamma M A \mmr {M'}
      \and
      \jl\mmr\Gamma {\gsubst{A'}{M'}} S \mmmr {S'} {A''}
    }{
      \jl\mr\Gamma {\prd x {A} {A'}} {\spinecons M S} \mmmr {\spinecons
      {M'}{S'}} {A''}
    }

    \infer{ }{
      \jl\mr\Gamma {\app{\cst a} S} \spinenil \mr \spinenil {{\app{\cst a} S}}
    }
  \end{mathpar}

  \fbox{$\js\mr\Gamma\sigma{\Gamma'}\mmr{\sigma'}$}
  \qquad
  {Substitution}

@@

    % Cons
    \infer{
      \js\mr\Gamma\sigma{\Gamma'}\mmr{\sigma'}
      \and
      \jm\mmr\Gamma M {\gsubst A {\sigma'}} \mmmr {M'}
    }{
      \js\mr\Gamma{(\msubstcons \sigma x M)} {(\envcons {\Gamma'} x {A'})}
      \mmmr {(\msubstcons {\sigma'} x {M'})}
    }

    % Nil
    \infer{ }{
      \js\mr\Gamma\msubstnil\envnil\mr\envnil
    }

@@

The main judgement is atomic object typing $\jf\mr\Gamma F \mmr {F'}
A$, which means that in repository $\mr$ and environment $\Gamma$, the
applicative term $F$ is well-typed of type $A$, and the slicing of $F$
produces a new repository $\mmr$. The object $F'$ is the
\emph{residual} of this operation: only certain kind of applications
are sliced, and it is possible that the ``tip'' of application $F$ is
not sliced (because it is declared as non-sliceable). In that case,
$M$ is this tip; if $F$ is directly sliceable, then $M$ is the
metavariable referring to it.

@@

\begin{figure*}
  \fbox{$\jsig\mr\Sigma\mmr{\Sigma'}$}
  \qquad
  {Signature}

@@

    \infer{
      \jsig\mr\Sigma\mmr{\Sigma'}
      \and
      \jfam\mmr\gnil {\!_{\Sigma'} A}\mmmr {A'}
    }{
      \jsig\mr{\gcons\Sigma{{\cst c :^L A}}}\mmmr{\gcons{\Sigma'}{{\cst c :^L A'}}}
    }

    \infer{
      \jsig\mmr\Sigma\mmmr{\Sigma'}
      \and
      \jkind\mr\gnil {\!_{\Sigma'} K}\mmr {K'}
    }{
      \jsig\mr{\gcons\Sigma{{\cst a : K}}}\mmmr{\gcons{\Sigma'}{{\cst a : K'}}}
    }

    \infer{ }{
      \jsig\mr\gnil\mr\gnil
    }
  \end{mathpar}

  \fbox{$\jfam\mr\Gamma A \mmr {A'}$}
  \qquad
  {Family}

@@

    \infer{
      \jfam\mr\Gamma {A_1} \mmr {A_1'}
      \and
      \jfam\mmr{\gcons\Gamma{\tp {\var x} {A_1'}}} {A_2} \mmmr {A_2'}
    }{
      \jfam\mr\Gamma {\prd x {A_1} {A_2}} \mmmr {\prd x {A_1'} {A_2'}}
    }

    \infer{
      \Sigma(a) = K
      \and
      \jfl\mr\Gamma K S \mmr {S'}
    }{
      \jfam\mr\Gamma {\app{\cst a} S} \mmr {\app{\cst a} {S'}}
    }
  \end{mathpar}

  \fbox{$\jfl\mr\Gamma K S \mmr {S'}$}
  \qquad
  {Family spine}

  \begin{mathpar}
    \infer{
      \jm\mr\Gamma M A \mmr {M'}
      \and
      \jfl\mmr\Gamma {\gsubst{K}{M'}} S \mmmr {S'}
    }{
      \jfl\mr\Gamma {\prd x {A} {K}} {\spinecons M S} \mmmr {\spinecons
      {M'}{S'}}
    }

    \infer{ }{
      \jfl\mr\Gamma\type\spinenil\mr\spinenil
    }
  \end{mathpar}

  \fbox{$\jkind\mr\Gamma K \mmr {K'}$}
  \qquad
  {Kind}

  \begin{mathpar}
    \infer{
      \jfam\mr\Gamma A\mmr{A'}
      \and
      \jkind\mmr{\envcons\Gamma x A} K \mmmr {K'}
    }{
      \jkind\mr\Gamma{\prd x A K}\mmmr{\prd x {A'}{K'}}
    }

    \infer{ }{\jkind\mr\Gamma\type\mr\type}
  \end{mathpar}

>>>>

\begin{align*}
  K &\gequal
  \prd x A K \gor
  \type &
  \text{Kind}\\
  A &\gequal
  \prd x A A \gor
  P &
  \text{Type family} \\
  P &\gequal
  \app {\cst a} S &
  \text{Atomic type} \\
  M &\gequal
  \lam x M \gor
  F &
  \text{Canonical object} \\
  F &\gequal \app H S
  \gor
  \smeta X \sigma &
  \text{Atomic object} \\
  H &\gequal
  \var x \gor
  \cst c &
  \text{Head}\\
  S &\gequal
  \spinenil \gor
  \spinecons M S &
  \text{Spine}\\
  \sigma &\gequal
  \msubstnil \gor
  \msubstcons \sigma x M &
  \text{Parallel substitution} \\
  \Gamma &\gequal
  \envnil \gor
  \envcons \Gamma x A &
  \text{Environment}
\end{align*}

@@

A \emph{repository} $\mr$ is a finite map from metavariable to
judgments of atomic object well-typing, along with a distinguished
metavariable, corresponding to the \emph{tip} of the term contained in
it:
$$ \mr\;:\;(\meta X \mapsto (\Gamma\vdash\tp F P)), \smeta X
\sigma $$
The algorithm will maintain the invariant that a repository
$\mr$ is \emph{acyclic}, that is for all $\meta X$ in $\mr$, the
fixpoint of $\FMV{\meta X}$ does not contain $\meta X$. If $\mr =
(\Delta, X[\sigma])$, we write $\hat {\mr}$ for $X[\sigma]$, abusively
$\mr$ for $\Delta$ and $\mr (\smeta X \sigma)$ for $\gsubst M\sigma$
if $\Delta(\meta X) = (\Gamma\vdash M : A)$.

@@

  \fbox{$\jm\Delta\Gamma M A {\Delta'} {M'}$}
  \qquad
  Canonical object

  \begin{mathpar}
    \infer[MLam]{
      \jm\Delta{\envcons\Gamma x A} M {A'} {\Delta'} {M'}
    }{
      \jm\Delta\Gamma {\lam x M} {\prd x A {A'}} {\Delta'} {\lam x M'}
    }

    \infer[MAtom]{
      \jf\Delta\Gamma F {\Delta'} {F'} {P}
       \and
      \jea{\Delta'}\Gamma P {\app{\cst a} S}
    }{
      \jm\Delta\Gamma F {\app{\cst a} S} {\Delta'} {F'}
    }
    % TODO commenter type ground
  \end{mathpar}

  \fbox{$\jf\Delta\Gamma F {\Delta'} {F'} P$}
  \qquad
  {Atomic object}

@@

    % Non-sliceable
    \infer[FAppNoSlice]{
      \jh\Delta\Gamma H A \annnonsliceable
      \and
      \jl\Delta\Gamma A S {\Delta'} {S'} P
    }{
      \jf\Delta\Gamma {\app H S} {\Delta'} {\app H {S'}} P
    }

    % Meta
    \infer[FMeta]{
      \Delta(\meta X) = (\Gamma'\vdash F:P)
      \and
      \js\Delta\Gamma \sigma {\Gamma'} {\Delta'} {\sigma'}
    }{
      \jf\Delta\Gamma {\smeta X \sigma} {\Delta'} {\smeta X {\sigma'}} {\gsubst{P}{\sigma'}}
    }

    % Sliceable
    \infer[FAppSlice]{
      \jh\Delta\Gamma H A \annsliceable
      \and
      \jl\Delta\Gamma A S {\Delta'} {S'} {P}
      \and
      \stren{\Gamma, \app H {S'}}=\Gamma'
    }{
      \jf\Delta\Gamma {\app H S} {\mapadd{\Delta'}{\meta X}{(\Gamma'\vdash
          {\app H {S'}}:P)}} {\smeta X{\msubstid{\Gamma'}}} P
    }
    \quad\text{($\meta X$ fresh in $\Delta'$)}

    % Defined
    \infer[FAppEval]{
      \jh\Delta\Gamma H A {\anndefined T}
      \and
      % check args pas nécessaire (fait ds la red)
      % \jl\Delta\Gamma A S {{\Delta'}} {S'} P
      % \and
      \jcc\Delta\Gamma A T S {\Delta''} F
      % check result pas nécessaire (fait ds la red)
      % \and
      % \jm{\Delta''}\Gamma F P {\Delta'''} {F'}
    }{
      \jf\Delta\Gamma {\app H S} {\Delta'''} {F'} P
    }

@@

  \fbox{$\jh\Delta\Gamma H A L$}
  \qquad
  {Head}

  \begin{mathpar}
    % constant
    \infer[HConst]{
      \Sigma(\cst c) = A; L
    }{
      \jh\Delta\Gamma {\cst c} A L
    }

    % variable
    \infer[HVar]{
      \Gamma(\var x) = A
    }{
      \jh\Delta\Gamma {\var x} A \annnonsliceable
    }
  \end{mathpar}

  \fbox{$\jl\Delta\Gamma A S {\Delta'}{S'} P$}
  \qquad
  {Spine}

  \begin{mathpar}
    \infer[SCons]{
      \jm\Delta\Gamma M A {\Delta'} {M'}
      \and
      \jl{\Delta'}\Gamma {\gsubst{A'}{\msubst x {M'}}} S {\Delta''} {S'} P
    }{
      \jl\Delta\Gamma {\prd x {A} {A'}} {\spinecons M S} {\Delta''} {\spinecons
      {M'}{S'}} P
    }

    \infer[SNil]{ }{
      \jl\Delta\Gamma {\app{\cst a} S} \spinenil \Delta \spinenil {{\app{\cst a} S}}
    }
  \end{mathpar}

  \fbox{$\js\Delta\Gamma\sigma{\Gamma'}{\Delta'}{\sigma'}$}
  \qquad
  {Substitution}

@@

    % Cons
    \infer[$\sigma$Cons]{
      \js\Delta\Gamma\sigma{\Gamma'}{\Delta'}{\sigma'}
      \and
      \jm{\Delta'}\Gamma M {\gsubst A {\sigma'}} {\Delta''} {M'}
    }{
      \js\Delta\Gamma{(\msubstcons \sigma x M)} {(\envcons {\Gamma'} x {A'})}
      {\Delta''} {(\msubstcons {\sigma'} x {M'})}
    }

    % Nil
    \infer[$\sigma$Nil]{ }{
      \js\Delta\Gamma\msubstnil\envnil\Delta\envnil
    }

@@

The main judgement is atomic object typing $\jf\Delta\Gamma F
{\Delta'} {F'} P$, which means that in repository $\mr$ and environment $\Gamma$, the
applicative term $F$ is well-typed of type $A$, and the slicing of $F$
produces a new repository $\mmr$. The object $F'$ is the
\emph{residual} of this operation: only certain kind of applications
are sliced, and it is possible that the ``tip'' of application $F$ is
not sliced (because it is declared as non-sliceable). In that case,
$M$ is this tip; if $F$ is directly sliceable, then $M$ is the
metavariable referring to it.

@@

\begin{figure*}
  \fbox{$\jsig\Delta\Sigma{\Delta'}{\Sigma'}$}
  \qquad
  {Signature}

@@

    \infer{
      \jsig\Delta\Sigma{\Delta'}{\Sigma'}
      \and
      \jfam{\Delta'}\gnil {\!_{\Sigma'} A}{\Delta''} {A'}
    }{
      \jsig\Delta{\gcons\Sigma{{\cst c :^L A}}}{\Delta''}{\gcons{\Sigma'}{{\cst c :^L A'}}}
    }

    \infer{
      \jsig{\Delta'}\Sigma{\Delta''}{\Sigma'}
      \and
      \jkind\Delta\gnil {\!_{\Sigma'} K}{\Delta'} {K'}
    }{
      \jsig\Delta{\gcons\Sigma{{\cst a : K}}}{\Delta''}{\gcons{\Sigma'}{{\cst a : K'}}}
    }

    \infer{ }{
      \jsig\Delta\gnil\Delta\gnil
    }
  \end{mathpar}

  \fbox{$\jfam\Delta\Gamma A {\Delta'} {A'}$}
  \qquad
  {Family}

@@

    \infer{
      \jfam\Delta\Gamma {A_1} {\Delta'} {A_1'}
      \and
      \jfam{\Delta'}{\gcons\Gamma{\tp {\var x} {A_1'}}} {A_2} {\Delta''} {A_2'}
    }{
      \jfam\Delta\Gamma {\prd x {A_1} {A_2}} {\Delta''} {\prd x {A_1'} {A_2'}}
    }

    \infer{
      \Sigma(a) = K
      \and
      \jfl\Delta\Gamma K S {\Delta'} {S'}
    }{
      \jfam\Delta\Gamma {\app{\cst a} S} {\Delta'} {\app{\cst a} {S'}}
    }
  \end{mathpar}

  \fbox{$\jfl\Delta\Gamma K S {\Delta'} {S'}$}
  \qquad
  {Family spine}

  \begin{mathpar}
    \infer{
      \jm\Delta\Gamma M A {\Delta'} {M'}
      \and
      \jfl{\Delta'}\Gamma {\gsubst{K}{\msubst x {M'}}} S {\Delta''} {S'}
    }{
      \jfl\Delta\Gamma {\prd x {A} {K}} {\spinecons M S} {\Delta''} {\spinecons
      {M'}{S'}}
    }

    \infer{ }{
      \jfl\Delta\Gamma\type\spinenil\Delta\spinenil
    }
  \end{mathpar}

  \fbox{$\jkind\Delta\Gamma K {\Delta'} {K'}$}
  \qquad
  {Kind}

  \begin{mathpar}
    \infer{
      \jfam\Delta\Gamma A{\Delta'}{A'}
      \and
      \jkind{\Delta'}{\envcons\Gamma x A} K {\Delta''} {K'}
    }{
      \jkind\Delta\Gamma{\prd x A K}{\Delta''}{\prd x {A'}{K'}}
    }

    \infer{ }{\jkind\Delta\Gamma\type\Delta\type}
  \end{mathpar}
